---
title: "coordinate_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{coordinate_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we will be covering the functions to transform between coordinate representations. We will start with the representation `a <- c(-10,0,10)` of a longitude or latitude.

```{r}
library(NcLibrary)

a <- c(-10,0,10)
lat_to_str(a)
lon_to_str(a)
```

Note that as default the 0 is given one of the coordinates. But 0W is equivalent to 0E and equally with the latitude. Now we will see how to do the reverse change.

```{r}

lon <- c("10W","0W","0E","10E")
lat <- c("10S","0S","0N","10N")

str_to_lon(lon)
str_to_lat(lat)
```

Finally the functions return error if it is not a good string representation.

```{r}

lon <- c("10N") # Wrong longitude
lat <- c("10N") # Wrong latitude

tryCatch({str_to_lon(lon)},error=function(e) print(e))

tryCatch({str_to_lat(lat)},error=function(e) print(e))
```

