---
title: "NcLibrary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NcLibrary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval= FALSE
)
```


## Introduction

The **NcLibrary** package provides functions for retrieving data from the ERA5 database directly to R. Additionally, it offers various functions for modifying existing lists.

To install NcLibrary, use the following command:

```{r setup, message=FALSE}
devtools::install_github("Davidjmz12/NcLibrary")
```

For a detailed summary of the library, refer to the [official documentation](https://davidjmz12.github.io/NcLibrary/).

## Setting Up the Environment

This section is meant for those interested in creating requests to ERA5 databases.  If this is not applicable to your needs, feel free to skip this section.

Firstly we must register on the [Climate Data Store](https://cds.climate.copernicus.eu/#!/home) to obtain an API key. 

Upon successful registration, proceed by following the steps outlined in the  *Install the CDS API key* section of the [official guide](https://cds.climate.copernicus.eu/api-how-to). In case the URL or key of the guide file fails to appear, consider refreshing the page while ensuring that you are logged in.

Finally, we have to set the *R* environment. As the execution of Python code is essential, the R library *reticulate* is employed. An independent *conda* environment with the libraries needed will be created after using the following function. Please note that if this configuration has been completed previously, it may be skipped.

``` r
set_reticulate()
```
If there is not any conda interpreter in your computer you could use, for example, the function `install_miniconda()`. 

If the *reticulate* environment is no longer needed, it can be deleted using the following function:

``` r
close_reticulate()
```

## Temperature in Spain at 2000

### Data Preparation

The **NcLibrary** contains a range of functions for obtaining desired information from the **ERA5 Complete Reanalysis Database**.  In this example, we will be extracting the daily temperature at 12:00 in the *Iberian Peninsula* with a grid resolution of 1x1 degrees. We select full levels at 171 hPa and 820 hPa.

Two methods are available for retrieving this data:

(a) Use the User Interface via `create_nc_file_app()`.

(b) Employ the function `create_nc_file_data()` to do it manually.

In our example, we will use the second method. However, it is crucial to have the IDs of the desired variables and levels (for more information, refer to the [variables](https://apps.ecmwf.int/codes/grib/param-db/) and [levels](https://confluence.ecmwf.int/display/UDOC/L137+model+level+definitions)) hyperlinks).

With all the information collected, we create a file "data/example.nc" using the previously mentioned function.
```{r, message=FALSE}
library(NcLibrary)
create_nc_file_data("71,112", "12:00:00", "1.0x1.0", "data/example.nc", "2000/01/01", "2000/12/31", "40/-15/30/15", "130")
```

Please be aware that this function can be relatively slow for substantial data requests. Once this function finishes,  a *.nc* file containing all the data will be created. 

### NetCDF file data extraction

After generating a .nc file, it is essential to access a preview of the information contained within. This enables efficient utilization of functions (e.g., selecting variable names). To achieve this, the `nc_open()` function of the `netcdf4` library can be utilized, returning a list with valuable insights regarding the .nc file.

```{r}
nc_open("data/example.nc")
```
### Data Formatting

The next step involves formatting the file.   This is accomplished using the `read_nc_file()` which returns a dataframe with the information of the file. In this example, we will create a basic configuration list to read our nc-file. For further details, please consult  `help("read_nc_file")`.

```{r}
configuration <- list(dim=list(
                                latitude=list(type="lat", coord_360=T, name="latitude", format=0),
                                longitude=list(type="lon", coord_360=T, name="longitude", format=0)
                              ),
                       time=list(extended=F,time_div=T,name="time"),
                       sep="_",
                       nc_file="data/example.nc")
spain_list <- read_nc_file(configuration)
```

## Observed temperature in Madrid

The previous section contains the main functions of the library. In addition to that, there are functions to extract data from **ECAD databases**.

### Data Preparation

In this case, the ECAD page already has an interface to request the data directly (see [here](https://www.ecad.eu/dailydata/customquery.php)). Therefore, the request must be done beforehand. In this example, we will imagine that we have a file called **data/madrid_t.zip** with the blended data of the station *Madrid-Barajas* in a period of time containing the year 2000.

### Format the data

Upon downloading the data, extraction can be performed using the `read_eca_zip()` function. It returns a list with the data of all the files within the **.zip** file. It is important to use the parameter `blended` to indicate whether our file is blended or not.

```{r}
madrid_obs <- read_eca_zip("data/madrid_t.zip",blended=TRUE)
```
