---
title: "Read NetCDF4 File"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read NetCDF4 File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we will see a few examples focused on how to use the `read_nc_file` function. Firstly the most usual usage of the function is without any further argument. This example will use a basic file already downloaded using the function `create_nc_file_data` and will display a subset of the data frame.

```{r setup}
library(NcLibrary)

listNC <- read_nc_file("data/one_test.nc")

listNC$df[c(1:6,40)]

```

This is the default configuration. However there are a lot of other features. For example if the .nc file has a very thin density which is not really needed, we can apply a reduction in both directions. By doing this, the resulting data frame will be reduced considerably.

```{r}

test_2_nr <- read_nc_file("data/other_test.nc")
test_2_r <- read_nc_file("data/other_test.nc", reduction.x=2L,reduction.y=2L)

cat("Data frame dimension with no reduction:", dim(test_2_nr$df), "\n")

cat("Data frame dimension with two times reduction in both directions:", dim(test_2_r$df), "\n")

```

Moreover a list of variables and levels can be selected from the whole file. To see which variables the .nc file holds we can use the function `fetch_information`.

```{r}
fetch_information("data/one_test.nc")$variables
fetch_information("data/one_test.nc")$geo_levels
```

Now with this information we can just choose a selection of them and the resulting list will only have this information.

```{r}
test_3 <- read_nc_file("data/one_test.nc",var.names="u",geo.levels = 10)
cat("Variables: ", test_3$vars, "\n")
cat("Levels: ", test_3$geo_levels, "\n")
```

Sometimes a full hour representation is needed. To work with extended hours is important to use the parameter `extended.hours=TRUE`. Moreover we can change the separator character in the column names by using `sep`. With this configuration keep in mind that ":" is reserved for hours and can make confusion if you use it as a separator.

```{r}
test_4 <- read_nc_file("data/one_test.nc",extended.hours=TRUE, sep="-")
cat("Previous example hours: ", test_3$hours, "\n")
cat("This example hours: ", test_4$hours, "\n")
cat("One column name: ", names(test_4$df)[5])
```

Finally as default, the hours are a new dimension. However, this behavior can be change using the parameter  `hourly.col=FALSE`.

```{r}
test_5 <- read_nc_file("data/one_test.nc",hourly.col=FALSE)
test_5$df[,1:6]
```

In this example we see that the hours are new row entries in the data frame. Note that this configuration combined with `extended.hours=TRUE` produces new columns with the seconds, minutes and the full extended hour:

```{r}
test_6 <- read_nc_file("data/one_test.nc",hourly.col=FALSE, extended.hours=TRUE)
test_6$df[,1:9]
```